---
title: "Heart_project"
output: html_document
---

**P&S Project on Data Analysis and Visualization**

For our project we decided to analyze the dataset from https://www.kaggle.com/fedesoriano/heart-failure-prediction

concerning the heart diseases and health parameters which might be indicators to the health disease

and used further to predict the possibility of a heart disease given these parameters.

We decided to focus on few of the parameters and show how age, cholesterol level and sex might correlates to heart diseases.



```{r}
library(ggplot2)
library(BSDA)
library(fitdistrplus)
```


### Reading and processing the data
```{r}
heart <- read.csv("heart.csv")
```

**General information about the data obtained**

```{r}

summary(heart)

```

**Visualizing the correlation between parameters and Heart Failures**

```{r}

ggplot(heart,aes(x=Age,fill=factor(HeartDisease)))+geom_density(alpha=0.5)+
 xlab(label = "Age")+
 ggtitle("Correlation between Heart Diseases and Age")

```


```{r}

ggplot(heart,aes(x=Cholesterol,fill=factor(HeartDisease)))+geom_density(alpha=0.5)+
 xlab(label = "Cholesterol")+
 ggtitle("Correlation between Heart Diseases and Cholesterol")

```

```{r}
ggplot(heart,aes(x=HeartDisease,fill=factor(Sex)))+geom_density(alpha=0.5)+
 xlab(label = "Heart Disease")+
 ggtitle("Correlation between Heart Diseases and Sex")

```
**However, let's take a look on the number of males and females**
```{r}
summary(heart$Sex)

summary(heart$Sex == "M")

```
Therefore, there is information about 193 females and 725 males in this dataset.

Even with this inequality in numbers between genders, we see that there are more females without

heart diseases, than males, whose number in the given statistics of the dataset is 3.5 times larger than females'


**Determination of the Age and Cholesterol distributions**

```{r}
descdist(heart$Age)

```

Given these results, we might assume that distribution of the age in the given statistics is close to normal.

Let's apply this to Cholesterol as well
```{r}
descdist(heart$Cholesterol)

```
Cholesterol distribution is close to normal as well, but it is not that close as Age's distribution. We'll see it in a bit.


### Testing

First, let's visualize the distribution of the Age and Cholesterol (without correlation to heart failure, as we did it before)

```{r}

cholesterol <- heart$Cholesterol
cholesterol = cholesterol[cholesterol != 0]
hist(heart$Age)
lines(x=dnorm(heart$Age, mean=mean(heart$Age), sd=sd(heart$Age)))
hist(cholesterol)
```
We see that Age and Cholesterol distributions look like normal distributions, though skewed.

Cholesterol distribution is right skewed.


Now, let's finally test if age or cholesterol data are normally distributed. For this we will use Shapiro-Wilk test for normality.

**Normal distribution**

```{r}
shapiro.test(heart$Age)
shapiro.test(cholesterol)

```
As we can see, p-values for both test converge to zero, so neither age nor cholesterol data follow normal distribution, even though p-value

of the Age is a lot bigger than cholesterol's, which could be seen from the graphics above.


Let's test it on exponential distribution using Kolmogorov-Smirnov test.

**Exponential distribution**


```{r}
ks.test(heart$Age, 'pexp', rate=1/mean(heart$Age))
ks.test(cholesterol, 'pexp', rate=1/mean(cholesterol))
```
So we see that age and cholesterol doesn't follow exponential distribuition either.

**Cholesterol and Age**


Now let's test if Age and Cholesterol level are independent using chi-squared test.

```{r}
chisq.test(heart$Cholesterol, heart$Age)
```
We see that p-value converges to zero, that means that age and cholesterol level are stronlgly correlated.

Now let's test other features concerning the Heart Disease

**Age and Heart Diseases**


And start with correlation between Age and Heart Diseases

```{r}
chisq.test(heart$Age, heart$HeartDisease)

```
p-value goes to zero therefore Heart Diseases depend on the Age of the person.

**Cholesterol and Heart Diseases**

Now the same for Cholesterol

```{r}
chisq.test(heart$Cholesterol, heart$HeartDisease)

```
p-value is close to zero therefore Heart Diseases depend on the Cholesterol level of the person.


**Age and Heart Diseases**


Now let's find out whether heart diseases depend on the sex of the person

```{r}
chisq.test(heart$Sex, heart$HeartDisease)

```

P-value converges to zero which supports the theory that Heart Diseases depend on the sex




### Conclusion

To conclude, we have shown the dependence of Age, Sex and Cholesterol level on the Heart Failure

and shown interesting correlations between them, especially unobvious fact about females having

less heart failures than males, and pretty obvious such that risk of the heart failure increases with the age.

We have learnt how to work with real-life data and apply knowledge obtained during the course

to a real problem. We were lucky that chosen dataset is very clean, without any NaN values and ready to use in

practice. It really helped us to focus on analyzing the dependencies between features and making assumptions about the 

distributions of the data. It turned out that real life data rarely follows the ideal patterns of distributions that we learned

but it is easy to approximate them to normal (even if they dont follow it) and work with them pretty easy.
